---
title: "Situations to match the flowchart"
output: 
  github_document:
    toc: true
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = FALSE)

library(ggplot2)
library(dplyr)

plot_sv <- function(scenario_df) {
  
  c_sum <- scenario_df %>%
    group_by(ts) %>%
    summarize(N = dplyr::n(),
              E = sum(energy),
              M = sum(mass),
              Ebar = mean(energy),
              Mbar = mean(mass)
    ) %>%
    ungroup() %>%
    tidyr::pivot_longer(-ts, names_to = "variable", values_to = "value") %>% 
    mutate(currency = substr(variable, 0, 1))
  
  #return(list(
  ggplot(c_sum, aes(x = ts, y = value, color = currency)) +
    geom_point() +
    geom_line() +
    scale_color_viridis_d(end = .8) +
    facet_wrap(vars(variable), scales = "free", as.table = F, nrow = 3) +
    expand_limits(y = 0) +
    theme_bw() +
    theme(legend.position = "none")#,
}

plot_spectra <- function(scenario_df) {
  
  gridExtra::grid.arrange(grobs = list(
    
    ggplot(scenario_df, aes(x = mass, weights = energy)) +
      geom_density() +
      theme_bw() +
      xlim(0, 1.5 * (max(scenario$mass)))+
      facet_wrap(vars(ts)) +
      ggtitle("E spectrum: nb individuals by size weighted by E"),
    ggplot(scenario_df, aes(x = mass)) +
      geom_density() +
      theme_bw() +
      xlim(0, 1.5 * (max(scenario$mass))) +
      facet_wrap(vars(ts)) +
      ggtitle("Size spectrum: nb individuals by size")
  ),
  nrow = 2)
  #))
}

sv_ratios <- function(scenario) {
  
  ratios = scenario %>% 
    group_by(ts) %>%
    summarize(E = sum(energy),
              N = dplyr::n(),
              M = sum(mass)) %>%
    ungroup()
  
  data.frame(
    variable = c("E ratio", "N ratio"),
    ratio = c(
      ratios$E[2]/ratios$E[1],
      ratios$N[2]/ratios$N[1]
    )
  )
  
}

```

# Beginning

Beginning with a community of 1000 energy units divided into two species of fairly different sizes. The two species have *equal shares of the energy*. This means there are *more individuals of the smaller species*.

```{r toy setup, fig.dim = c(4,4)}

E = 1000

M1 = 5
M2 = 15

E1 = M1 ^ .75
E2 = M2 ^ .75

N1 = ceiling((E/2)/E1)
N2 = ceiling((E/2)/E2)

N = N1 + N2

S = 2

individuals_t1 <- data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 1
)

gridExtra::grid.arrange(grobs = list(
  
  ggplot(individuals_t1, aes(x = mass, weights = energy)) +
    geom_density() +
    theme_bw() +
    xlim(0, 1.5 * (max(individuals_t1$mass)))+
    facet_wrap(vars(ts)) +
    ggtitle("E spectrum: nb individuals by size weighted by E"),
  ggplot(individuals_t1, aes(x = mass)) +
    geom_density() +
    theme_bw() +
    xlim(0, 1.5 * (max(individuals_t1$mass))) +
    facet_wrap(vars(ts)) +
    ggtitle("Size spectrum: nb individuals by size")),
  nrow = 2)

```

# E is stable

## Stable E, N decreases

Achieved by increasing the average size.

Average metabolic rate increases.

The energy spectrum becomes skewed large.

The size spectrum becomes more large-skewed than it was, but note that *beginning with energetic equivalence means beginning with a small-skewed size spectrum*.

```{r E stable N decrease, fig.dim = c(4,4)}


N1 = ceiling((.25 * E)/E1)
N2 = ceiling((.75 * E)/E2)

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)
```

## E is stable, N increases

Achieved by decreasing the average size.

Average metabolic rate decreases.

The size spectrum in individuals becomes more concentrated at small sizes, and the energy spectrum skews small. 

```{r E stable N increase, fig.dim = c(4,4)}


N1 = ceiling((.75 * E)/E1)
N2 = ceiling((.25 * E)/E2)

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)
```

# E increases 

## E increases and N increases proportionally

Here we have increased E by 1.5 by amplifying the original community without changing the underlying size structure.

```{r E increase, N increase proportional, fig.dim = c(4,4)}

Eamp = E * 1.5

N1 = ceiling((Eamp/2)/E1)
N2 = ceiling((Eamp/2)/E2)


scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)
```

## E increases, N increases but not as much as E

Gains went to large species more than to small species.

```{r E increase, N increases less, fig.dim = c(4,4)}


N1 = ceiling((E/2)/E1)
N2 = ceiling((E/2)/E2)

N1 = ceiling(1.2 * N1)
N2 = ceiling(1.75 * N2)


scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```

## E increases, N increases more than E

Gains go disproportionately to small species.

### Additional energy available only to small species
```{r E increase for small, fig.dim = c(4,4)}

N1 = ceiling(((E/2) + (.2 * E))/E1)
N2 = ceiling((E/2)/E2)


scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```

### E increase available to all sizes

```{r E increase for all, fig.dim = c(4,4)}


N1 = ceiling((E/2 + 500)/E1)
N2 = ceiling((E/2 + 500)/E2)


scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```

## E increases, N does not change
Achieved by reallocating small individuals to large individuals.

The spectra shift large.

Average energy use and size increase.

```{r E increase, N stable, fig.dim = c(4,4)}


N1 = ceiling((E/2)/E1)
N2 = ceiling((E/2)/E2)

N2 = N2 + ceiling(.75 * N1)
N1 = ceiling(.25 * N1)


scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```

## E increases, N decreases

Small species decline, and large species increase. The greater the size difference the easier it is to get this pattern; if they're too close it might actually be impossible to get it very convincingly, because E scales at .75. 


```{r E increase, N decreases, fig.dim = c(4,4)}


N1 = ceiling((E/2)/E1)
N2 = ceiling((E/2)/E2)

N2 = N2 + ceiling(.4 * N1)
N1 = ceiling(.5 * N1)

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```


# E decreases

## E decreases and N increases proportionally

Resources decrease for all sizes, turning-down-the-volume on the whole community without changing the size structure.

```{r E decreases, N proportionally, fig.dim = c(4,4)}

Edamp = .75 * E

N1 = ceiling((Edamp/2)/E1)
N2 = ceiling((Edamp/2)/E2)

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```

## E decreases and N decreases more than E

Small species decline. Large species can increase (to a limit) or decline, but loss of energy to large species must be less than to small species. 

```{r E decreases, N decreases more, fig.dim = c(4,4)}


N1 = ceiling(.75 * ((E/2)/E1))
N2 = ceiling((E/2) / E2)

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```

## E decreases and N decreases less than E

Large species decline disproportionately.

```{r E decreases, N decreases less, fig.dim = c(4,4)}


N1 = ceiling(((E/2)/E1))
N2 = ceiling(.75 * ((E/2)/E2))

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```

## E decreases and N is stable

Large species decline disproportionately.

Average e must decrease. 

```{r E decreases, N stable, fig.dim = c(4,4)}


N1 = ceiling(((E/2)/E1))
N2 = ceiling(((E/2)/E2))

N1 = N1 + ceiling(.25 * N2)
N2 = ceiling(.75 * N2)

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```


## E decreases and N increases

Large species *must decline* and small species *must gain*.

This can only result in a *decrease* in E to a threshold/with a particular arrangement of relative sizes and the increase. 


```{r E decreases, N increases, fig.dim = c(4,4)}


N1 = ceiling(((E/2)/E1))
N2 = ceiling(((E/2)/E2))

N1 = ceiling(N1 * 1.5)
N2 = ceiling(.3 * N2)

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```