---
title: "Situations to match the flowchart"
output: 
  html_document:
    toc: true
---

```{r setup, echo = F}
knitr::opts_chunk$set(echo = FALSE)

library(ggplot2)
library(dplyr)

plot_sv <- function(scenario_df) {
  
  c_sum <- scenario_df %>%
    group_by(ts) %>%
    summarize(N = dplyr::n(),
              E = sum(energy),
              M = sum(mass),
              Ebar = mean(energy),
              Mbar = mean(mass)
    ) %>%
    ungroup() %>%
    tidyr::pivot_longer(-ts, names_to = "variable", values_to = "value") %>% 
    mutate(currency = substr(variable, 0, 1))
  
  #return(list(
  ggplot(c_sum, aes(x = ts, y = value, color = currency)) +
    geom_point() +
    geom_line() +
    scale_color_viridis_d(end = .8) +
    facet_wrap(vars(variable), scales = "free", as.table = F, nrow = 3) +
    expand_limits(y = 0) +
    theme_bw() +
    theme(legend.position = "none")#,
}

plot_spectra <- function(scenario_df) {
  
  gridExtra::grid.arrange(grobs = list(
    
    ggplot(scenario_df, aes(x = mass, weights = energy)) +
      geom_density() +
      theme_bw() +
      xlim(0, 1.5 * (max(scenario$mass)))+
      facet_wrap(vars(ts)) +
      ggtitle("E spectrum: nb individuals by size weighted by E"),
    ggplot(scenario_df, aes(x = mass)) +
      geom_density() +
      theme_bw() +
      xlim(0, 1.5 * (max(scenario$mass))) +
      facet_wrap(vars(ts)) +
      ggtitle("Size spectrum: nb individuals by size")
  ),
  nrow = 2)
  #))
}

sv_ratios <- function(scenario) {
  
  ratios = scenario %>% 
    group_by(ts) %>%
    summarize(E = sum(energy),
              N = dplyr::n(),
              M = sum(mass)) %>%
    ungroup()
  
  data.frame(
    variable = c("E ratio", "N ratio", "M ratio", "avg e change", "avg size change", "size change / e change"),
    ratio = c(
      ratios$E[2]/ratios$E[1],
      ratios$N[2]/ratios$N[1],
      ratios$M[2]/ratios$M[1],
      (ratios$E[2]/ratios$E[1]) / (ratios$N[2]/ratios$N[1]),
      (ratios$M[2] / ratios$N[2]) / (ratios$M[1] / ratios$N[1]),
      (ratios$M[2]/ratios$M[1]) /   (ratios$E[2]/ratios$E[1])
    )
  )
  
}

```

# Beginning

Beginning with a community of 1000 energy units divided into two species of fairly different sizes. The two species have *equal shares of the energy*. This means there are *more individuals of the smaller species*.

```{r toy setup, fig.dim = c(4,4)}

E = 1000

M1 = 5
M2 = 15

E1 = M1 ^ .75
E2 = M2 ^ .75

N1 = ceiling((E/2)/E1)
N2 = ceiling((E/2)/E2)

N = N1 + N2

S = 2

individuals_t1 <- data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 1
)

gridExtra::grid.arrange(grobs = list(
  
  ggplot(individuals_t1, aes(x = mass, weights = energy)) +
    geom_density() +
    theme_bw() +
    xlim(0, 1.5 * (max(individuals_t1$mass)))+
    facet_wrap(vars(ts)) +
    ggtitle("E spectrum: nb individuals by size weighted by E"),
  ggplot(individuals_t1, aes(x = mass)) +
    geom_density() +
    theme_bw() +
    xlim(0, 1.5 * (max(individuals_t1$mass))) +
    facet_wrap(vars(ts)) +
    ggtitle("Size spectrum: nb individuals by size")),
  nrow = 2)

```

# E is stable

## Stable E and N

### Completely stable
The trivial version of this is, nothing about anything changes. 

```{r E stable N stable, fig.dim = c(4,4)}

scenario <-  bind_rows(individuals_t1, 
                       mutate(individuals_t1, ts = 2))

plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)
```

But (as I belatedly realized) we can in fact achieve stable N and E any number of ways, by shifting size distribution. N and E only constrain mean e. 

### Size spectrum consolidates

So we can converge more to the mean:

```{r E stable N stable converging, fig.dim =c(4,4)}

mean_e = sum(E1 * N1, E2 * N2) / sum(N1, N2)

mean_mass = mean_e ^ (4/3)

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = mean_e,
  mass =mean_mass,
  ts = 2
))


plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```

In which case, the total M and the average size both decrease. 

### Size spectrum polarizes

Or the size spectrum can become more polarized. 

In which case the total M and the average size *increase* slightly.

```{r E stable N stable polarizing, fig.dim =c(4,4)}

mean_e = sum(E1 * N1, E2 * N2) / sum(N1, N2)

N1 = 164
N2 = 52

E1 = .75 * E1
M1 = E1 ^ (4/3)

E2 = ((mean_e * (N1 + N2)) - (E1 * N1)) / N2
M2 = E2 ^ (4/3)

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))


plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```

```{r E stable N stable M stable but change, fig.dim =c(4,4)}

E = 1000

M1 = 5
M2 = 15

E1 = M1 ^ .75
E2 = M2 ^ .75

N1 = ceiling((E/2)/E1)
N2 = ceiling((E/2)/E2)

E = (N1 * E1) + (N2 * E2)

M1 = 2
# M2 = 15

E1 = M1 ^ .75
# E2 = M2 ^ .75

N1 = 100

N2 = N - N1

E2 = (E - (E1 * N1))/N2
M2 = E2 ^ (4/3)

S = 2

scenario <-  bind_rows(individuals_t1, data.frame(
  species = c(rep(1, times = N1),
              rep(2, times = N2)),
  energy = c(rep(E1, times = N1),
             rep(E2, times = N2)),
  mass =c(rep(M1, times = N1),
          rep(M2, times = N2)),
  ts = 2
))


plot_spectra(scenario)
plot_sv(scenario)
sv_ratios(scenario)

```


My intuition is you cannot get E, N, and M *all* stable and have any meaningful change to the size spectrum, but my attempts to demonstrate this analytically have been a nightmare, and I'm leaning towards "don't put anything past highly squishy combinations of numbers". 

Option B is to try and leverage the overlap metric?


<!-- ```{r E stable N decrease, fig.dim = c(4,4)} -->


<!-- N1 = ceiling((.25 * E)/E1) -->
<!-- N2 = ceiling((.75 * E)/E2) -->

<!-- scenario <-  bind_rows(individuals_t1, data.frame( -->
<!--   species = c(rep(1, times = N1), -->
<!--               rep(2, times = N2)), -->
<!--   energy = c(rep(E1, times = N1), -->
<!--              rep(E2, times = N2)), -->
<!--   mass =c(rep(M1, times = N1), -->
<!--           rep(M2, times = N2)), -->
<!--   ts = 2 -->
<!-- )) -->

<!-- plot_spectra(scenario) -->
<!-- plot_sv(scenario) -->
<!-- sv_ratios(scenario) -->
<!-- ``` -->
